### 2017-01-30 visualizemicrosteps.Rmd
code to make microstep plots

---
title: "Visualizing `RSiena`'s Microsteps"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

## Fit a model to data

I'll start like I always do: with the friends data. This code is directly from the [`RSiena` webpage](https://www.stats.ox.ac.uk/~snijders/siena/Rscript02SienaVariableFormat.R).

```{r get_data, cache=TRUE}
library(RSiena)
setwd("data/s50_data")
friend.data.w1 <- as.matrix(read.table("s50-network1.dat"))
friend.data.w2 <- as.matrix(read.table("s50-network2.dat"))
friend.data.w3 <- as.matrix(read.table("s50-network3.dat"))
drink <- as.matrix(read.table("s50-alcohol.dat"))
smoke <- as.matrix(read.table("s50-smoke.dat"))
friendship <- sienaDependent(
                     array(c(friend.data.w1, friend.data.w2,
                             friend.data.w3),
                     dim = c(50, 50, 3)))
smoke1 <- coCovar(smoke[ , 1 ])
alcohol <- varCovar(drink)
mydata <- sienaDataCreate(friendship, smoke1, alcohol)
myeffnull <- getEffects(mydata)
myeffalt1 <- includeEffects(myeffnull, transTrip)
myalgorithm <- sienaAlgorithmCreate(projname = 's50_3')
set.seed(823746)
ansnull <- siena07(myalgorithm, data = mydata, effects = myeffnull,
	                 returnChains = TRUE, returnDataFrame = TRUE,
	                 returnDeps = TRUE, silent = TRUE, verbose = FALSE,
	                 batch = TRUE)
ansalt1 <- siena07(myalgorithm, data = mydata, effects = myeffalt1,
                    returnChains = TRUE, returnDataFrame = TRUE,
                    returnDeps = TRUE,silent = TRUE, verbose = FALSE,
                    batch = TRUE)
```

Look at a plot of the data:

```{r plot}
library(geomnet)
w1 <- fortify(as.adjmat(friend.data.w1))
w2 <- fortify(as.adjmat(friend.data.w2))
w3 <- fortify(as.adjmat(friend.data.w3))
id <- na.omit(unique(c(w1$from, w1$to, w2$from, w2$to, w3$from, w3$to)))
nodes <- data.frame(id = rep(id[order(parse_number(id))], 3),
                    wave = rep(1:3, each = length(id)),
                    drink = c(drink[-c(13,20),1],drink[-c(13,20),2],
                              drink[-c(13,20),3]), 
                    smoke = rep(smoke[-c(13,20),1],3),
                    stringsAsFactors = F)
w1$wave <- 1
w2$wave <- 2
w3$wave <- 3
fnet <- rbind(w1,w2,w3)
fnet <- merge(fnet, nodes, by.x = c('from', 'wave'),
              by.y = c('id',"wave"), all = T)
# need to change the id to be 0-49 to match the chains coding
# library(dplyr)
# fnet %>% 
 # mutate(from = parse_number(from) - 1, 
  #       to = parse_number(to) - 1) -> fnet
library(readr)

ggplot(data = fnet) + 
  geom_net(aes(from_id = from, to_id = to), fiteach = T) + 
  facet_wrap(~wave, labeller = 'label_both', nrow = 1)
```

What do the fitted models look like?

```{r fits}
summary(ansnull)
summary(ansalt1)
```

Convergence looks good in both cases!

Now, I want to visualize the individual changes, starting with wave 1 to wave 2 from the null model.

```{r getchains, cache=TRUE}
get_chain_info <- function(ans){
  L <- length(ans$chain)
  M <- length(ans$chain[[1]][[1]])
  bigres <- NULL
  res <- data.frame()
  for (l in 1:L){
    res <- data.frame()
    for (m in 1:M){
      sub <- data.frame(plyr::ldply(
          lapply(ansnull$chain[[l]][[1]][[m]], unlist),
        rbind), stringsAsFactors = FALSE)
      sub$rep <- l
      sub$period <- m
      res <- rbind(res, sub)
    }
    bigres[[l]] <- res  
  }
  ret <- plyr::rbind.fill(bigres)
  return(ret)
}
ansnullchains <- get_chain_info(ansnull)
```

```{r firstwave}
library(dplyr)
library(readr)
ansnullchains %>% 
  filter(period == 1) %>% 
  group_by(rep) %>%
  select(rep, from = X4, to = X5) %>% 
  mutate(val = as.numeric(!from == to),
         from = paste0("V", parse_number(from)+1), # make the chains
         to = paste0("V", parse_number(to)+1) # match the data vars 
         ) -> ansnullchainsw1w2

addMicrostep <- function(dat, newedge){
  N <- nrow(dat)
  selfie <- newedge$from == newedge$to
  check <- dplyr::filter(dat, from == newedge$from & to == newedge$to)
  if (nrow(check) == 0 && !selfie){
    dat2 <- tibble::add_row(dat, from = newedge$from, to = newedge$to)
    dat2$col <- FALSE
    dat2$col[N+1] <- TRUE # color the new edge. 
  } else if (nrow(check) == 1 && !selfie){
    dat2 <- dplyr::filter(dat, !(from == newedge$from & to == newedge$to))
    dat2$col <- FALSE
  } else{
    dat2 <- dat
    dat2$col <- FALSE
  }
  return(dat2)
}

listMicrosteps <- function(dat, microsteps){
  iters <- NULL
  iters[[1]] <- dat
  B <- nrow(microsteps)
  for (i in 2:(B+1)){
    iters[[i]] <- addMicrostep(iters[[(i-1)]], microsteps[(i-1),])
  }
  return(iters)
}

# test it on one of the reps from the null model

ms1 <- listMicrosteps(dat = w1, 
                      microsteps = filter(ansnullchainsw1w2, rep == 1))
ms1[[200]]
# looks legit
```

Now, I want to plot all of them! Make a movie!

```{r movie}
# layout the original wave
# get (x,y) coordinates to add to the edge lists
# merge nodes and microsteps before plotting.

library(sna)
w1.net <- as.network(friend.data.w1)
layoutxy <- gplot.layout.fruchtermanreingold(friend.data.w1, layout.par = NULL)
layoutxy <- data.frame(layoutxy)
names(layoutxy) <- c("x", "y")
layoutxy$id <- get.vertex.attribute(w1.net, "vertex.names")

w1_nodes_xy <- merge(nodes %>% filter(wave == 1), layoutxy, by = "id", all = T)

plotms <- function(dat, nodes){
  dat2 <- merge(dat, nodes, by.x = "from", by.y= "id", all=T)
  p <- ggplot(data = dat2) + 
    geom_net(aes(from_id = from, to_id = to, x = x, y = y),
             layout.alg = NULL, size = 1.5, directed = T, linewidth = .5,
             arrow = arrow(angle = 20, 
                           length = unit(5, "pt"), type = "open"),
             arrowgap = .02) + #, ecolour = c("grey40", "red")[as.factor(dat2$col)]) + 
    theme_net()
  return(p)
}

setwd("plots")
for (i in 2:length(ms1)) {
  # creating a name for each plot file with leading zeros
  print(i)
  if (i < 10) {name = paste('plots/000',i,'plot.png',sep='')}
  if (i < 100 && i >= 10) {name = paste('plots/00',i,'plot.png', sep='')}
  if (i >= 100) {name = paste('plots/0', i,'plot.png', sep='')}
  png(name)
  print(plotms(ms1[[i]], nodes = w1_nodes_xy))
  dev.off()
} 
make.mov <- function(){
     unlink("plot.mpg")
     system("convert -delay 0.5 plot*.jpg plot.mpg")
}
make.mov()
```